---
title: "Homework 3"
author: "Stefan Konigorski"
date: "November 2, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Compute frequencies in the Pima diabetes dataset

Load the Pima diabetes dataset:

```{r}
load(file = url("https://www.dropbox.com/s/hxn1c7yw9r8aw8v/Pima_diabetes.RData?dl=1"))
#head(Pima_diabetes)

```

Which variables are measured on a nominal level?
Answer: Outcome.

Now compute frequency tables, barplots, and mosaic plots of all nominal variables in the dataset.
```{r}
# absolute frequencies
table(Pima_diabetes$Outcome)

#relative frequencies
library(janitor)
janitor::tabyl(Pima_diabetes$Outcome)

#cummulative frequencies
n <- nrow(Pima_diabetes)
data <- table(Pima_diabetes$Outcome)/n
cum_fre <-cumsum(data)
cum_fre

#generating plots
#barplots
barplot(table(Pima_diabetes$Outcome))

#mosaic plot
da <- table(Pima_diabetes$Age, Pima_diabetes$Outcome)
mosaicplot(da, main = "Showing Outcome in terms of Age",
           color = "skyblue2",
           xlab = "Age",
           ylab = "Outcome",
           border = "chocolate")
```

Next, create a variable which describes whether a woman had more or less than 4 pregnancies. Then, use this variable to create a 2x2 table with diabetes outcome. Do you see an indication of whether the number of pregnancies is associated with diabetes prevalence? Do you think your investigation is a good way to investigate this?

```{r}
Pima_diabetes$pre <- ifelse(Pima_diabetes$Pregnancies < 4, 0, 1)
table(Pima_diabetes$pre, Pima_diabetes$Outcome)
```
According to the table, women who have been pregnant more than four times are more likely to have diabetes than women who have had fewer pregnancies. Though pregnancy has a strong link to diabetes, I believe that other parameters such as BMI, Glucose, and Blood Pressure are important factors to consider when determining diabetes prevalence.


## Exercise 2: Generate a table with descriptive statistics (optional, but recommended)

Use any dataset (a dataset that you have worked with in the past, or that you are currently working with, a dataset that is available on Blackboard, in R or that you have downloaded from the internet), and generate a table with descriptive statistics of the main variables of interest.

## Exercise 3: Plots using ggplot2

Load the NoShow dataset:

```{r}
load(file = url("https://www.dropbox.com/s/4oqg79cn1qfnhsh/NoShowdata.RData?dl=1"))
head(NoShowdata)
library(ggplot2)
#boxplot showing age stratified by neighborhood.
ggplot(NoShowdata, aes(Age, Neighbourhood)) + geom_boxplot()
# Histogram
ggplot(NoShowdata, aes(Age)) + geom_histogram()

#Create a histogram of Age, stratified by whether the person showed up - in one panel using the or in multiple panels.

ggplot(NoShowdata, aes(Age)) +
  geom_histogram(aes(color = `No-show`), fill = "white",
                position = "identity", bins = 30) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"))
#Stratify this plot further by gender.
ggplot(NoShowdata, aes(Age, fill= `No-show`)) +
  geom_histogram(aes(color = Gender),alpha=0.2,position = "identity", bins = 50)
```

Use ggplot2 to generate the following plots:

- Create a boxplots of Age (stratified) by neighborhood.
- Create a histogram of Age.
- Create a histogram of Age, stratified by whether the person showed up - in one panel using the or in multiple panels.
- Stratify this plot further by gender.

What insights can you get from these plots? For which goal would you create these plots?

Answer:

People between the ages of 20 and 50 are more likely to miss appointments. The number increases as they get older. Females are more absent than men in the gender base.
