---
title: "Homework 7"
author: "Stefan Konigorski"
date: "November 30, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Preparations

Load KiGGS data:

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
kiggs <- KiGGS03_06
```

## Exercise 1: Multiple imputation

a) In the KiGGS dataset, choose 2 metric variables of your choice.
```{r}
w <- kiggs$E014z #Weight gain during pregnancy (kg)
n <- kiggs$e029z #Nights in the Hospital (EFB)

#Weight gain during pregnancy (kg)
str(w)
summary(w)
hist(as.numeric(as.character(w)))

##Nights in the Hospital (EFB)
str(n)
summary(n)
hist(as.numeric(as.character(n)))
```





b) Compute the correlation between these two variables
```{r}
# Preparing dataset:
data_complete <- data.frame(weight_gain = as.numeric(as.character(w)), 
                           night_hospital = as.numeric(as.character(n)))

idx_complete <- complete.cases(data_complete)
data_complete <- data_complete[idx_complete, ]

table(!is.na(data_complete$weight_gain))
table(!is.na(data_complete$night_hospital))

cor(data_complete$weight_gain, data_complete$night_hospital)

res <- cor.test(data_complete$weight_gain, data_complete$night_hospital, method = "pearson")
res$estimate
res$conf.int
res$p.value


# visualization
library("ggpubr")
ggscatter(data_complete, x = "weight_gain", y = "night_hospital", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson")

```



c) Perform a multiple imputation using the mice function with the 'pmm' method and otherwise default settings.
```{r}
library(mice)

# Pattern of missing values
md.pattern(data_complete)

tempData <- mice(data_complete, m = 5, maxit = 5, meth = 'pmm', seed = 500)
summary(tempData)

```


d) Compute the correlation again on the multiple imputed datasets.
```{r}
#importing library
library(miceadds)
miceadds::micombine.cor(mi.res = tempData, variables = c(1,2), conf.level = 0.95, method = "pearson")
```